Model Checking Differentially Private Properties

### 2018-01-27 mdp.tex
more changes ;)

>>>>

\paragraph{Model Checking Algorithm.}

Our next task is to compute $\{ M, \neighbor{S}, s \models
\dpriv{\epsilon}{\delta} \phi \}$ for a Markov decision process $M =
(S, \Act, \wp, L)$. Recall the semantics of $\dpriv{\epsilon}{\delta}
\phi$. Given $s, t$ with $s \neighbor{M} t$ and a path formula
$\phi$, we need to decide whether
$\myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq
\epsilon \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} + \delta$
for every scheduler $\scheduler{S}$.
If the path formula only contain next operators, it is easy to deal with. For $\phi:=\X B$ with $B\subseteq S$ only one step behaviour needs to be considered, thus only the first action in the query sequence is needed to compute the maximum
\begin{eqnarray}
  \label{eqn:max-nextB}
\max_{\scheduler{S}}
\myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\gX B} -
\epsilon \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\gX B}
\end{eqnarray}
with $s, t \in S$. It can be easily generalized to nested next operators:  one needs only to enumerate 
all actions query sequences of length obtained by the number of nexted $\X$ operators.

In general, the computation of the maximal probability for MDPs is non-trivial. \lz{we argue that it is difficult and leave as future work}

Finally, we the property 
$\wedge_{i\in N} \dpriv{\epsilon}{0}(\X^k \top) \wedge \dpriv{\epsilon}{0}(F \top)$ that is used to express the privacy for our threshold Markov chains. It is only a valid formula if we would allow infinite conjunctions. However, it is interesting to note the relation to language equivalence for probabilistic automata\lz{to be discussed}

### 2018-01-27 semantics.tex
more changes ;)

>>>>

The model checking algorithm for $\dpCTLstar$ can be treated as in the classical setting \cite{BK:08:PMC}, as all we need is to  compute the the probability
  $\myPr{s}{K}{\neighbor{S}}{\phi}$ with general path formula $\phi$, one first construct a deterministic $\omega$-automaton $R$, such as deterministic Rabin automaton. Then, the probability reduces to a 
  reachability probability in the product Markov chain obtained from $K$ and $R$. There are more efficient algorithms without the product construction, see \cite{}. We skip the details.

### 2018-01-27 specifications.tex
more changes ;)

Further, consider the Markov chain in
Figure~\ref{figure:geometric-mechanism-markov-chain}. Assume
$L(out_i) = \{ out_i \}$ and $\Delta (f) = 1$. Then $in_k$ and $in_l$ are
neighbors iff $| k - l | \leq 1$\lz{put above sentence to the def of the model}. Define the $\dpCTL$ formula
$\psi = \dpriv{.5}{0} (\F out_0) \wedge \dpriv{.5}{0} (\F out_1) \wedge
\cdots \wedge \dpriv{.5}{0} (\F out_5)$\lz{I think we better use $\X$ operator here ;)}. If the state $in_k$ satisfies
$\psi$ for $k = 0, \ldots, 5$, then the $.5$-geometric mechanism is
$(0.5, 0)$-differentially private.

@@

\paragraph{Tighter Privacy Bounds for Compositional Properties}
One of the advantage of applying model checking is that we may get tighter bounds for compositional properties. Consider our survey example, and the property $\dpriv{1}{.5} (\X Y)$. Obviously, it holds in states $+$ and $-$. A careful check ensures that one cannot decrease the value of $\epsilon_1=1$ or $\delta_1=.5$ without increasing the other one.

@@

>>>>

 Define the $\dpCTL$ formula
$\psi = \dpriv{.5}{0} (\F out_0) \wedge \dpriv{.5}{0} (\F out_1) \wedge
\cdots \wedge \dpriv{.5}{0} (\F out_5)$\lz{I think we better use $\X$ operator here ;)}. If the state $in_k$ satisfies
$\psi$ for $k = 0, \ldots, 5$, then the $.5$-geometric mechanism is
$(0.5, 0)$-differentially private.

@@

\paragraph{Tighter Privacy Bounds for Compositional Properties.}
One of the advantage of applying model checking is that we may get tighter bounds for compositional properties. Consider our survey example, and the property $\dpriv{1}{.5} (\X Y)$. Obviously, it holds in states $+$ and $-$. A careful check ensures that one cannot decrease the value of $\epsilon_1=1$ or $\delta_1=.5$ without increasing the other one.

@@

\paragraph{Reachability Privacy} 
For our threshold \lz{to do} model, we have the desired privacy property $\wedge_{i\in N} \dpriv{\epsilon}{0}(\X^k \top) \wedge \dpriv{\epsilon}{0}(F \top)$. Here $\dpriv{\epsilon}{0}(F \top)$ is a reachability privacy: it states that related states should have similar probabilities of reaching $\top$.
